### 2025-12-09 dataPrep.Rmd
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP

---
title: "dataPrep"
output: html_document
---

```{r long classification dataset - paper}

library(tidyverse)
library(countrycode)

"%ni%" = Negate("%in%")

datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")

datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")

datasetCountries <- datasetCountries %>% 
  dplyr::select(ID, countryPartialF) %>% 
  rename(country = countryPartialF)

datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
                                        by = "ID") %>% 
  dplyr::select(-reader)

datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)

table(datasetClassificationCountries$country)

# 1) split into one row per country
datasetClassificationCountries_long <- datasetClassificationCountries %>%
  separate_rows(country, sep = ",\\s*") %>%
  mutate(countries = str_trim(country))

datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
                                                               datasetClassificationCountries_long$countries))
# fix country names 
# 1. Direct mapping
datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia")) # virtually filtering for papers that would not have a matching country

# add the year
datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
  dplyr::select(ID, Year)

datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
                                                   datasetYear,
                                                   by = "ID", all.x = T, all.y = F)

df_cat_long <- datasetClassificationCountries_long_clean %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag, -None)

write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)

```

```{r long classification dataset - survey}

library(tidyverse)
library(countrycode)

"%ni%" = Negate("%in%")

surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")

surveyClassification <- surveyClassification %>% 
  filter(Inclusion == 1)

surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
                                                 surveyClassification$surveyedCountries)

# 1) split into one row per country
surveyCountries_long <- surveyClassification %>%
  separate_rows(surveyedCountries, sep = ";\\s*") %>%
  mutate(countries = str_trim(surveyedCountries))

# fix country names first manual then with iso3

surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))

# 1. Direct mapping
surveyCountries_long_clean <- surveyCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))

surveyCountries_long_clean_yearly <- surveyCountries_long_clean
surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)

surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
  # 1. build a single “years” string
  mutate(
    all_years = if_else(
      is.na(repeatedMeasure) | repeatedMeasure == "",
      as.character(year),
      paste(year, repeatedMeasure, sep = ";")
    ),
    # drop every space character
    all_years = str_remove_all(all_years, "\\s+")
  ) %>%
  separate_rows(all_years, sep = ";") %>%
  mutate(year = as.integer(all_years)) %>%
  dplyr::select(-all_years, -repeatedMeasure)

df_cat_long <- surveyCountries_long_clean_yearly_t %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  dplyr::select(-flag)

write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)

```

```{r putting data together}

library(tidyverse)
"%ni%" = Negate("%in%")

datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")

datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")

datasetCountries <- datasetCountries %>% 
  select(ID, countryPartialF) %>% 
  rename(country = countryPartialF)

datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
                                        by = "ID") %>% 
  select(-reader)

datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)

table(datasetClassificationCountries$country)

# 1) split into one row per country
datasetClassificationCountries_long <- datasetClassificationCountries %>%
  separate_rows(country, sep = ",\\s*") %>%
  mutate(countries = str_trim(country))

datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
                                                               datasetClassificationCountries_long$countries))
# fix country names 
library(countrycode)

# 1. Direct mapping
datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))

df_cat_long <- datasetClassificationCountries_long_clean %>%
  # keep only “paper belongs to this category”
  filter(sum(Perceptions, Priority, Policy, Health, Behavior) >= 1,
         None != 1)

# 2) get total papers by country
papers_by_country <- df_cat_long %>%
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_papers") %>%
  arrange(desc(n_papers))

datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")

datasetPollution_avg <- datasetPollution %>% 
  filter(Year %in% 2019) %>%
  group_by(Region) %>% 
  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))

datasetPollution2 <- read.csv("dataNew/GlobalPM25-V5GL0502-Annual-1998-2023-wThresFrac.csv")

datasetPollution2_avg <- datasetPollution2 %>% 
  filter(Year %in% 2019) %>%
  filter(Region %in% c("Hong Kong", "Taiwan")) %>% 
  group_by(Region) %>% 
  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))

datasetPollution_avg <- rbind(datasetPollution_avg, datasetPollution2_avg)

datasetPollution_ctr <- datasetPollution_avg %>%
  mutate(
    iso3 = countrycode(Region,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(!is.na(iso3)) %>% 
  rename(countryName = Region)

papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)

fullCountrylist <- unique(countrycode::codelist$iso3c)

papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)

## adding survey nb

surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")

surveyClassification <- surveyClassification %>% 
  filter(Inclusion == 1)

surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
                                                 surveyClassification$surveyedCountries)

# 1) split into one row per country
surveyCountries_long <- surveyClassification %>%
  separate_rows(surveyedCountries, sep = ";\\s*") %>%
  mutate(countries = str_trim(surveyedCountries))

# fix country names first manual then with iso3

surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))

# 1. Direct mapping
surveyCountries_long_clean <- surveyCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))

surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2017 - 2021", 2019, surveyCountries_long_clean$year)
surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2021 - 2022", 2021, surveyCountries_long_clean$year)

surveyCountries_long_clean_t <- surveyCountries_long_clean %>%
  # 1. build a single “years” string
  mutate(
    all_years = if_else(
      is.na(repeatedMeasure) | repeatedMeasure == "",
      as.character(year),
      paste(year, repeatedMeasure, sep = ";")
    ),
    # drop every space character
    all_years = str_remove_all(all_years, "\\s+")
  ) %>%
  separate_rows(all_years, sep = ";") %>%
  mutate(year = as.integer(all_years)) # %>%
  # dplyr::select(-all_years, -repeatedMeasure)

# 2) get total surveys by country
surveys_by_country <- surveyCountries_long_clean_t %>%
  filter(Inclusion == 1) %>% 
  group_by(iso3, .drop = F) %>% 
  count(iso3, name = "n_surveys") %>%
  arrange(desc(n_surveys))

papers_by_country_pollution_survey <- merge(papers_by_country_pollution, surveys_by_country, by = "iso3", all.x = T, all.y = T)

papers_by_country_pollution_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_survey$n_surveys), 0, papers_by_country_pollution_survey$n_surveys)

population <- read.csv("dataNew/GHS_pop.csv", sep = ";")

popCtr <- population %>%
  group_by(GADM_ISO) %>% 
  summarize(population2015 = sum(X2015))

summaryDataset <- merge(papers_by_country_pollution_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)

summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
                  labels = c("very low", "low", "medium", "high", "very high"), 
                  right = FALSE)

summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)

### add income groups to the dataset

incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)

summaryDataset <- merge(summaryDataset, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)

plot(summaryDataset %>% 
       select(Income.group, pm25Avg, n_papers, n_surveys, population2015, categoryPollution))

## add region from natural earth

library(rnaturalearth)
library(terra)

world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))

world_countries_df <- as.data.frame(world_countries)

world_countries_df$iso_a3 <- ifelse(world_countries_df$iso_a3_eh == "FRA", "FRA",
                                    ifelse(world_countries_df$iso_a3_eh == "NOR", "NOR", world_countries_df$iso_a3))

summaryDataset2 <- merge(summaryDataset, world_countries_df %>% 
                          select(continent, region_un, iso_a3), by.x = "iso3", by.y = "iso_a3", all.x = T, all.y = F)

summaryDataset2 <- summaryDataset2 %>% 
  filter(!(iso3 == "FRA" & continent == "Seven seas (open ocean)")) %>% 
  filter(!(iso3 == "CYP" & countryName == "Northern Cyprus"))

write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)

```