### 2025-12-17 draftFigure1.Rmd
updated code and figure but need to review fig2 and update exposure fig2 heatmap

---
title: "draftFigure1"
output: html_document
---

```{r panels}

library(tidyverse)

dataset <- read.csv("output/datasetPlots.csv")

dataset$isDiff <- ifelse(dataset$isIndia == 1, "India",
                         ifelse(dataset$isChina == 1, "China", "Rest"))

summaryPaperInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_papers, na.rm = T)) %>% 
  filter(!is.na(Income.group))

summaryPaperInc$Income.group <- factor(
  summaryPaperInc$Income.group,
  levels = rev(c(
    "High income",
    "Upper middle income",
    "Lower middle income",
    "Low income"
  )),
  ordered = TRUE
)

# simple version of the plot
# summaryPaperInc %>% ggplot(aes(x = Income.group, y = nTotal))+
#   geom_bar(stat = "identity", aes(fill = isDiff))+
#   scale_fill_manual(values = c("#C19EAB", "#C19EAB", "#8B475D"))+
#   theme_minimal()+
#   labs(y = "Total number of papers", x = "Income group")+
#   theme(legend.position = "none")

totals <- summaryPaperInc %>%
  group_by(Income.group) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p1 <- summaryPaperInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#C19EAB", "#C19EAB", "#8B475D")) +
  theme_minimal() +
  labs(y = "Total number of papers", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p1

pdf("plots/Figure1_papers_by_income.pdf", width = 10)
p1
dev.off()

## then surveys
summarySurveysInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(Income.group))

summarySurveysInc$Income.group <- factor(
  summarySurveysInc$Income.group,
  levels = rev(c(
    "High income",
    "Upper middle income",
    "Lower middle income",
    "Low income"
  )),
  ordered = TRUE
)

totals <- summarySurveysInc %>%
  group_by(Income.group) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p2

pdf("plots/Figure1_surveys_by_income.pdf", width = 10)
p2
dev.off()

## then papers per exposure

summaryPaperExp <- dataset %>% 
  group_by(categoryPollution, isDiff) %>% 
  summarize(nTotal = sum(n_papers, na.rm = T)) %>% 
  filter(!is.na(categoryPollution))

summaryPaperExp$categoryPollution <- ifelse(
  !is.na(summaryPaperExp$categoryPollution),
  paste0(toupper(substr(summaryPaperExp$categoryPollution, 1, 1)),
         substr(summaryPaperExp$categoryPollution, 2, nchar(summaryPaperExp$categoryPollution))),
  NA
)

summaryPaperExp$categoryPollution <- factor(
  summaryPaperExp$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

totals <- summaryPaperExp %>%
  group_by(categoryPollution) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p3 <- summaryPaperExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#C19EAB", "#A67384", "#8B475D")) +
  theme_minimal() +
  labs(y = "Total number of papers", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p3

pdf("plots/Figure1_papers_by_exposure.pdf", width = 10)
p3
dev.off()

## then surveys per exposure

summarySurveysExp <- dataset %>% 
  group_by(categoryPollution, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(categoryPollution))

summarySurveysExp$categoryPollution <- ifelse(
  !is.na(summarySurveysExp$categoryPollution),
  paste0(toupper(substr(summarySurveysExp$categoryPollution, 1, 1)),
         substr(summarySurveysExp$categoryPollution, 2, nchar(summarySurveysExp$categoryPollution))),
  NA
)

summarySurveysExp$categoryPollution <- factor(
  summarySurveysExp$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

totals <- summarySurveysExp %>%
  group_by(categoryPollution) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p4 <- summarySurveysExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#D4ABEF", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p4

pdf("plots/Figure1_surveys_by_exposure.pdf", width = 10)
p4
dev.off()

```

```{r panels with eurobarometer highlight}
# globally there are 2390 survey items from eurobarometer out of 3140 total

library(tidyverse)
dataset <- read.csv("output/datasetPlots_eurobarometer.csv")

dataset <- dataset %>% 
  mutate(isDiff = ifelse(eurobarometerSource == 1, "Eurobarometer", "Rest"))

summarySurveysInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(Income.group))

summarySurveysInc$Income.group <- factor(
  summarySurveysInc$Income.group,
  levels = rev(c(
    "High income",
    "Upper middle income",
    "Lower middle income",
    "Low income"
  )),
  ordered = TRUE
)

totals <- summarySurveysInc %>%
  group_by(Income.group) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p2

pdf("plots/Figure1_surveys_by_income_eurob.pdf", width = 10)
p2
dev.off()

## then surveys per exposure

summarySurveysExp <- dataset %>% 
  group_by(categoryPollution, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(categoryPollution))

summarySurveysExp$categoryPollution <- ifelse(
  !is.na(summarySurveysExp$categoryPollution),
  paste0(toupper(substr(summarySurveysExp$categoryPollution, 1, 1)),
         substr(summarySurveysExp$categoryPollution, 2, nchar(summarySurveysExp$categoryPollution))),
  NA
)

summarySurveysExp$categoryPollution <- factor(
  summarySurveysExp$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

totals <- summarySurveysExp %>%
  group_by(categoryPollution) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p4 <- summarySurveysExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p4

pdf("plots/Figure1_surveys_by_exposure_eurob.pdf", width = 10)
p4
dev.off()

```

```{r maps}

library(tidyverse)
library(rnaturalearth)
library(terra)
library(tidyterra)
library(scales)          # for label formatting

dataset <- read.csv("output/datasetPlots.csv")

world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))

world_countries$iso_a3 <- ifelse(world_countries$iso_a3_eh == "FRA", "FRA",
                                    ifelse(world_countries$iso_a3_eh == "NOR", "NOR", world_countries$iso_a3))

world_countries_merged <- merge(world_countries, dataset
                                , by.x = "iso_a3", by.y = "iso3", all.x = T, all.y = F)

world_countries_merged$n_papers <- ifelse(is.na(world_countries_merged$n_papers), 0, world_countries_merged$n_papers)

my_breaks = c(1, 5, 20, 50, 150)

world_no_antarctica <- 
  world_countries_merged %>%
  filter(name != "Antarctica")

# 2) Build the map with integer labels and no decimals:
map_plot_clean <- ggplot() +
  geom_spatvector(
    data = world_no_antarctica,
    aes(fill = n_papers),
    color = "grey80",
    size  = 0.1
  ) +
  scale_fill_gradientn(
    colours = c("#DCC9D2", "#8B475D", "#4F2A30"),
    name    = "Papers",
    trans   = "log10",
    breaks  = my_breaks,
    labels  = label_number(
      accuracy = 1,
      big.mark = ",",
      trim     = TRUE
    ),
    na.value = "grey90"
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title    = "Global Publication Counts by Country",
    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5)
  )

# Print it
map_plot_clean

ggsave(plot = map_plot_clean, filename = paste("plots/mapPapersCount.png", sep = ""),
       dpi=600, width = 16, height = 12, units='cm')

pdf("plots/Figure1_mapPapersCount.pdf", width = 10)
map_plot_clean
dev.off()

## then for surveys
world_countries_merged$n_surveys <- ifelse(is.na(world_countries_merged$n_surveys), 0, world_countries_merged$n_surveys)

my_breaks = c(1, 5, 10, 30, 110)

# 1) Filter out Antarctica from your data:
world_no_antarctica <- 
  world_countries_merged %>%
  filter(name != "Antarctica")

# 2) Build the map with integer labels and no decimals:
map_plot_clean_survey <- ggplot() +
  geom_spatvector(
    data = world_no_antarctica,
    aes(fill = n_surveys),
    color = "grey70",
    size  = 0.1
  ) +
  scale_fill_gradientn(
    colours = c("#F3EBF8", "#D4ABEF", "#541F78"),
    name    = "Survey items",
    trans   = "log10",
    breaks  = my_breaks,
    labels  = label_number(
      accuracy = 1,
      big.mark = ",",
      trim     = TRUE
    ),
    na.value = "grey85"
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title    = "Global Survey Question Counts by Country",
    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5)
  )

# Print it
map_plot_clean_survey

ggsave(plot = map_plot_clean_survey, filename = paste("plots/mapSurveyCount.png", sep = ""),
       dpi=600, width = 16, height = 12, units='cm')

pdf("plots/Figure1_mapSurveyCount.pdf", width = 10)
map_plot_clean_survey
dev.off()

```